---
title: "Superstore with R"
author: "Virgil Chung Siu Tung"
date: '`r Sys.Date()`'
output: html_notebook
---

# Tableau Superstore Dataset

With growing demands and cut-throat competitions in the market, a Superstore Giant is seeking your knowledge in understanding what works best for them. They would like to understand which products, regions, categories and customer segments they should target or avoid.

They also want to have a Regression model to predict Sales or Profit.

# Get the dataset

Firstly, let's get the dataset from XLS file. In the XLS file, there are 3 sheets:

-   Orders: List of transactions
-   Returns: List of items returned
-   People: List of sales person for West, East, Central and South

Importing necessary libraries
```{r}
library(readxl)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(forcats)
library(scales)
```

Import the excel file into a RStudio as a dataframe
```{r}
df_order = read_xls('superstore.xls', sheet = 'Orders', col_names = TRUE)
df_return = read_xls('superstore.xls', sheet='Returns', col_names = TRUE)
df_people = read_xls('superstore.xls', sheet='People', col_names = TRUE)
```

Viewing data

"Order" dataframe:
```{r}
head(df_order)
```

"Return" dataframe:
```{r}
head(df_return)
```

"People" dataframe:
```{r}
head(df_people)
```

# Validate the data - "Order" data

Checking the structure of the "Order" data:

```{r}
str(df_order)
```

The "Order" dataframe is 9994 instances and 21 features. Now check the summary of the "Order" dataframe.

```{r}
summary(df_order)
```

What piqued interest was the statistics in Sales, Quantity, Discount and Profit column. From there, we know that:

-   Sales range from USD 0.44 to USD 22,638.48 per transaction; the average sales closed is USD 229.86 while a typical sale closed is USD 54.49
-   Quantity sold range from 1 to 14 items per transaction; typically each transaction sold around 3 to 4 items.
-   Discount range from USD 0 to USD 0.80, meaning this shop only give minimal discounts to customers.
-   Profit range from a loss USD -6,599.98 to a profit of USD 8,399.98. The average profit per transaction is USD 28.66 but we know most of the profit is lower than that due to median lower than mean.
-   On top of that, we also noticed that R is not able to identify which one is datetime format as they are parsed as characters.

Next, we check for missing values:

```{r}
apply(is.na(df_order), 2, sum)
```
There are no missing values for `df_order`. 

Next, we check on the unique values of each relevant features.
```{r, echo=FALSE}
cat(length(unique(df_order$Segment)), 'unique values in Segment: ', unique(df_order$Segment), '\n')
cat(length(unique(df_order$`Ship Mode`)), 'unique values in Ship Mode: ', unique(df_order$`Ship Mode`), '\n')
cat(length(unique(df_order$Country)), 'unique values in Country: ', unique(df_order$Country), '\n')
cat(length(unique(df_order$City)), 'unique values in City', '\n')
cat(length(unique(df_order$State)), 'unique values in State: ', unique(df_order$State), '\n')
cat(length(unique(df_order$Region)), 'unique values in Region: ', unique(df_order$Region), '\n')
cat(length(unique(df_order$Category)), 'unique values in Category: ', unique(df_order$Category), '\n')
cat(length(unique(df_order$`Sub-Category`)), 'unique values in Sub-Category: ', unique(df_order$`Sub-Category`), '\n')
```
Repeat the same for `df_return`. No need for `df_people` because it is just 4 rows.


# Validate the data - "Return" data

Structure of `df_return`:

```{r}
str(df_return)
```

Summary of `df_return`:

```{r}
summary(df_return)
```

Missing values of `df_return`:

```{r}
apply(is.na(df_return), 2, sum)
```

No missing values too in `df_return`.

Unique values of `df_return`:

```{r}
cat(length(unique(df_return$Returned)), 'unique values in Returned:', unique(df_return$Returned))
```

# Data Wrangling - Merging "Return" and "People" dataframe into "Order" dataframe

We merge `df_return$Returned` and `df_people$Person` into `df_order$Returned` and `df_order$Person` respectively, using full join to retain all values and rows.

```{r}
df = dplyr::full_join(df_order, df_return, by="Order ID")

df$`Order Date` = as.Date.character(df$`Order Date`, format="%Y-%m-%d")
df$`Ship Date` = as.Date(df$`Ship Date`, format="%Y-%m-%d")
df
```

`df` will be the dataset we use for EDA, transformation and prediction.

# Exploratory Data Analysis

In this section, we will drill deeper into the data for more insights. But first, we need to know what problems we want to solve and what questions to ask. Taking the POV of the owner of the Superstore:

-   Overview
    -   What is the monthly sales and profit since inception?
-   Overview - Increase Revenue
    -   Which product category and subcategory has the highest sales, with and without discount?
    -   Which customer segment that contribute to the highest sales?
    -   Which region, state and city contribute to the highest sales?
-   Overview - Reduce Loss
    -   Which product category and subcategory that has the highest returned item?
-   Prediction
    -   Regression: What is the overall sales and profit in the next month?
    -   Classification: Based on order features, what is the most likely ship mode for that particular order?

## Overview - Increase Revenue

###1. Which product category and subcategory with highest sales, with and without discount?

```{r}
df_category_sales <- df %>%
  select(Category, Sales) %>%
  group_by(Category) %>%
  summarise(Sales = sum(Sales))
```

```{r}
df_category_sales <- df_category_sales %>%
  mutate(Percentage = percent(df_category_sales$Sales/sum(df_category_sales$Sales)))
```

Pie Chart for Category by Sales Breakdown

```{r}
ggplot(df_category_sales, aes(x = "", y = Sales, fill = Category)) +
  ggtitle("Category by Sales Breakdown") +
  geom_col() +
   geom_label(aes(label = Percentage),
             position = position_stack(vjust = 0.5),
             show.legend = FALSE) +
  coord_polar(theta = "y")
```

Bar Chart for Category by Sales Breakdown

```{r}
bar_category_sales <- ggplot(data=df_category_sales, aes(x=Category, y=Sales, fill=Category)) +
  coord_cartesian(ylim = c(700000, 850000))+
  ggtitle("Category by Sales Breakdown") +
  geom_bar(stat="identity")+
  geom_text(aes(label=Percentage), vjust=1.6, color="white", size=3.5)+
  theme_minimal()
bar_category_sales + scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))
bar_category_sales
```

###2. Which customer segment that contribute to the highest sales?

```{r}
df_segment_sales <- df %>%
  select(Segment, Sales) %>%
  group_by(Segment) %>%
  summarise(Sales = sum(Sales))
```

Inputting percentage column into the table
```{r}
df_segment_sales <- df_segment_sales %>%
  mutate(Percentage = percent(df_segment_sales$Sales/sum(df_category_sales$Sales)))
```

Pie Chart for CUstomer Segment Sales Contribution
```{r}
ggplot(df_segment_sales, aes(x = "", y = Sales, fill = Segment)) +
  ggtitle("Customer Segment Sales Contribution") +
  geom_col() +
   geom_label(aes(label = Percentage ),
             position = position_stack(vjust = 0.5),
             show.legend = FALSE) +
  coord_polar(theta = "y")
```
###3. Which region, state and city contribute to the highest sales?

Region - Sales
```{r}
df_region_sales <- df %>%
  select(Region, Sales) %>%
  group_by(Region) %>%
  summarise(Sales = sum(Sales))
```

Inputting percentage column into the table
```{r}
df_region_sales <- df_region_sales %>%
    mutate(Percentage = percent(df_region_sales$Sales/sum(df_region_sales$Sales)))
```

Bar Chart for Region by Sales
```{r}
bar_region_sales <- ggplot(data=df_region_sales, aes(x=Region, y=Sales, fill=Region)) +
  coord_cartesian(ylim = c(300000, 800000))+
  ggtitle("Region by Sales Breakdown") +
  geom_bar(stat="identity")+
  geom_text(aes(label=Percentage), vjust=1.6, color="white", size=3.5)+
  theme_minimal()
bar_region_sales + scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9","#f5ad42"))
bar_region_sales
```

State - Sales
```{r}
df_state_sales <- df %>%
  select(State, Sales) %>%
  group_by(State) %>%
  summarise(Sales = sum(Sales))

df_state_sales <- df_state_sales[order(-df_state_sales$Sales),]
```

```{r}
df_state_sales <- df_state_sales %>%
  mutate(Percentage = percent(df_state_sales$Sales/sum(df_state_sales$Sales),accuracy = 0.01))
```

Horizontal Bar Chart for Top 10 Regions by Sales
```{r}
bar_state_sales <- ggplot(data=df_state_sales[1:10,], aes(x=State, y=Sales, fill=State)) +
  #coord_cartesian(ylim = c(300000, 800000))+
  geom_bar(stat="identity")+
  ggtitle("Top 10 Regions by Sales Breakdown") +
  geom_text(aes(label=Percentage), hjust=1.3, vjust=0.4, color="white", size=3)+
  theme_minimal()+
  coord_flip()
#bar_region_sales + scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9","#f5ad42"))
bar_state_sales
```

City - Sales 
```{r}
df_city_sales <- df %>%
  select(City, Sales) %>%
  group_by(City) %>%
  summarise(Sales = sum(Sales))

df_city_sales <- df_city_sales[order(-df_city_sales$Sales),]
```

```{r}
df_city_sales <- df_city_sales %>%
  mutate(Percentage = percent(df_city_sales$Sales/sum(df_city_sales$Sales),accuracy = 0.01))
```
Horizontal Bar Chart for Top 10 City by Sales
```{r}
bar_city_sales <- ggplot(data=df_city_sales[1:10,], aes(x=City, y=Sales, fill=City)) +
  geom_bar(stat="identity")+
  ggtitle("Sales by Category Breakdown") +
  geom_text(aes(label=Percentage), hjust=1.3, vjust=0.4, color="white", size=3)+
  theme_minimal()+
  coord_flip()

bar_city_sales
```








Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
